---
title: "Step 1: Data summary into DAs"
output: html_notebook
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

**Step 1: Data summary into DAs** This section takes each of the
environmental exposures and summarizes them into DA-level values. This
step must be redone for each city of interest.

**Packages** All packages are referenced using the :: command.
Therefore, they just need to be installed for the code to work. The only
package that needs to be turned on manually prior to the code being run
is tidyverse

List of packages needed: 
- tidyverse 
- sf 
- raster (version 3.1-5; devtools::install_version("raster", version = "3.1-5")) 
- sp 
- foreign 
- zoo

**Common steps** Each section basically works in the same way: a
function for the exposure type is called it, and you provide it paths to
the files needed to complete that section. There is then a bit of code
to check that it ran (visually inspect the output) and there is a list
of the attribute (columns) created.

**Save Locations and Naming Conventions** The code is set up to be saved
in an individual folder for each city. The exposure metric is placed in
the name of the file that is saved. Example:
"/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/cityname/cityname_exposure.shp"

Summarize buildings

```{r}
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/building.R")

# run for each city
buildings <- building(path_to_DA = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/DAs/Milton_DAs_Final/Milton_DAs_Final.shp",
                     path_to_building = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Buildings/Ontario_fixed.shp",
                     city_name = "milton",
                     save_file_path = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/milton/milton_building.shp")

# check map
buildings %>% 
    ggplot() +
    geom_sf(aes(fill= as.numeric(DA_area))) # note: could be "building_density" or "bldfn_d"

#DA_area = area of the DA (m^2)
#bldng_c = number of builidngs per DA
#bldng_d = building density (m^2)

# repeat for each city

```

Summarize NDVI

```{r}
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/NDVI.R")

ndvi <- NDVI(raster = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Landsat/NDVI/milton_ndvi.tif",
             census = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/DAs/Milton_DAs_Final/Milton_DAs_Final.shp",
             shapefile_save_location = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/milton/milton_NDVI.shp")

# check map
ndvi %>% 
    ggplot() +
    geom_sf(aes(fill= as.numeric(X2014))) 

# X2014 = average NDVI values for the year 2014
# X2015 = average NDVI values for the year 2015
# X2016 = average NDVI values for the year 2016
# X2017 = average NDVI values for the year 2017
# X2018 = average NDVI values for the year 2018

# repeat for each city
```

Summarize NO2

```{r}
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/NO2.R")

NO2 <- no2(raster = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/NO2/Canada_NO2_LUR_2014_16/Canada_AnnNO2_2014_16.tif",
         census = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/DAs/Milton_DAs_Final/Milton_DAs_Final.shp",
         shapefile_save_loc = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/milton/milton_no2.shp")

# check map
NO2 %>% 
    ggplot() +
    geom_sf(aes(fill= as.numeric(no2))) 

# no2 = averge 3-year concentration for nitrogen dioxide

# repeat for each city
```

Summarize PM2.5

```{r}
# set up to get PM2.5 files in order
# only need to do this once
pm <- raster::stack("/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/PM25/V4NA03_PM25_NA_201401_201412-RH35-NoNegs.asc/V4NA03_PM25_NA_201401_201412-RH35-NoNegs.asc")

list_file <- c("/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/PM25/V4NA03_PM25_NA_201401_201412-RH35-NoNegs.asc/V4NA03_PM25_NA_201401_201412-RH35-NoNegs.asc",
               "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/PM25/V4NA03_PM25_NA_201501_201512-RH35-NoNegs.asc/V4NA03_PM25_NA_201501_201512-RH35-NoNegs.asc",
               "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/PM25/V4NA03_PM25_NA_201601_201612-RH35-NoNegs.asc/V4NA03_PM25_NA_201601_201612-RH35-NoNegs.asc",
              "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/PM25/V4NA03_PM25_NA_201701_201712-RH35-NoNegs.asc/V4NA03_PM25_NA_201701_201712-RH35-NoNegs.asc",
              "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/PM25/V4NA03_PM25_NA_201801_201812-RH35-NoNegs.asc/V4NA03_PM25_NA_201801_201812-RH35-NoNegs.asc") 

# combine
pm <- raster::stack(list_file)
# rename
names(pm) <- c("X2014", "X2015", "X2016", "X2017", "X2018")

# save files
raster::writeRaster(pm, "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/PM25/combined_pm_2014_2018.tif")

# Run on all DAs, need to repeat for each city
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/pm2.5.R")
pm <- PM2.5(raster = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/PM25/combined_pm_2014_2018.tif",
            census = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/DAs/montrealmetro2.shp",
            shapefile_save_location = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/montreal/montrealmetro_pm2.5.shp")

# check map
pm %>% 
    ggplot() +
    geom_sf(aes(fill= as.numeric(X2014))) 

# X2014 = average PM2.5 concentration for the year 2014
# X2015 = average PM2.5 concentration for the year 2015
# X2016 = average PM2.5 concentration for the year 2016
# X2017 = average PM2.5 concentration for the year 2017
# X2018 = average PM2.5 concentration for the year 2018

x <- pm[rowSums(is.na(pm)) >0,]
# repeat for each city
```

Summarize power plants

```{r}
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/powerplant.R")

power_plant <- powerplant(vector = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/WRI_Powerplants/canada_powerplants.shp",
                         census = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/DAs/Milton_DAs_Final/Milton_DAs_Final.shp",
                         city_name = "milton",
                         shapefile_save_loc = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/milton/milton_powerplant.shp")

# check map
power_plant %>% 
    ggplot() +
    geom_sf(aes(fill= as.numeric(nearest_pp))) 

# nearest_pp = distance to nearest power plant

# repeat for each city
```

Summarize roads

```{r}
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/road.R")

# reading in road layer will save time in the processing
road_layer <- sf::read_sf("/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Roads/lrnf000r16a_e/lrnf000r16a_e.shp")

ROAD <- road(vector = road_layer,
             census = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/DAs/Milton_DAs_Final/Milton_DAs_Final.shp",
             city_name = "milton",
             shapefile_save_loc = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/milton/milton_roads.shp")

# check map
ROAD %>% 
    ggplot() +
    geom_sf(aes(fill= as.numeric(road_length))) 

# rd_lngt = road_length = length of all class 1,2,3,4 roads in each DA
# num_nds = num_nds = number of intersection of all class 1,2,3,4 roads in each DA

# repeat for each city
```

Summarize UV

```{r}
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/uv.R")

UV <- uv(raster = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/UV/m_d_do_sl.tif",
         census = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/DAs/Winnipeg_DAs_Final/Winnipeg_DAs_Final.shp",
         shapefile_save_loc = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/winnipeg/winnipeg_uv.shp")

# check map
UV %>% 
    ggplot() +
    geom_sf(aes(fill= as.numeric(m_d_do_sl))) 

# m_daily_do_sl = m_dly_d_s = Mean daily D dose sea level
# m_daily_do_alt = m_dly_d_l = mean daily D dose alt
# m_nn_d_sl = Mean noon D index sea level
# m_nn_d_alt = Mean noon D index alt
# m_95_nn_i_sl = m_95_nn__sl = 95th noon D index sea level
# m_95_nn_i_alt = m_95_nn__l = 95th noon D index alt

# repeat for each city
```

Summarize water

```{r}
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/water.R")

# read in water files first because it goes faster this way
fresh <- sf::read_sf("/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/water/ghy_000c11a_e/ghy_000c11a_e.shp")
sea <- sf::read_sf("/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/water/ghy_000h11a_e/ghy_000h11a_e.shp")

# only use for  Victoria
#fresh <- sf::read_sf("/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/water/fresh_dissolve.shp")
#sea <- sf::read_sf("/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/water/sea_dissolve.shp")



H2O <- water(path_to_DA = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/DAs/Windsor_DAs_Final/Windsor_DAs_Final.shp",
      fresh = fresh,
      sea = sea,
      city_name = "windsor",
      save_file_path = "/Users/zoedavis/Documents/untitled folder/out/windsor_water.shp")
        #"/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/windsor/windsor_water2.shp")



#H2O <- sf::read_sf("/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/ottawa/ottawa_water.shp")
# check map
water %>% 
    ggplot() +
    geom_sf(aes(fill= as.numeric(water_dist))) 

# water_dist = distance (m) to nearest water body
# water_area = area (m^2) of water in each DA

# repeat for each city

```


Summarize temperature - heat

```{r}
# This section allows you to select the days to search in GEE to download 
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/select_weather_and_dates.R")

# download weather data
weather <- weather_stations(city_name = "Vancouver",
                              prov = "BC")

#identify humid data to search for GEE data; days in which the humidex was above 30 or 35... can see what days could work
humid_days <- humid(weather_data = weather)

# calculate yearly temperature metrics
temp_stats <- mean_temperature_values(weather_data = weather) |> 
  mutate(year = lubridate::year(date)) |> 
    filter(mean_max_temp_3day > 30 & 
           mean_min_temp_3day >= 20) |> 
  group_by(year) |> 
  summarise(temp = n_distinct(date))




# Extract mean temperature for each DA after downloading Landsat thermal data from GEE
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/temperature_difference.R")

temp <-  temp_difference(raster = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/temperature/chicoutimi_temp.tif",
                         census = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/DAs/Chicoutimi_DAs_Final/Chicoutimi_DAs_Final.shp",
                         shapefile_save_location =
                          #"/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/scratch/milton_temp.shp"
                           "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/chicoutimi/chicoutimi_temp.shp"
                           ) 




temp %>% 
    ggplot() +
    geom_sf(aes(fill= as.numeric(d_DAtemp_mean))) +
  scale_fill_gradientn(colors = terrain.colors(20)) +
  labs(fill = "difference in temp")


# After you read it back in... column names
# temprtr = average temeprature within each DA
# men_tmp = the overall average temperature across the region
# d_DAtm_ = the difference in temperatrue between the DA mean and the mean of the region

```


Summarize temperature - cold
```{r}
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/select_weather_and_dates.R")

# download weather data
weather <- weather_stations(city_name = "Kelowna",
                              prov = "BC")

# identify cold days
cold_days <- cold(weather)
               

# Extract mean temperature for each DA after downloading Landsat thermal data from GEE
source("/Users/zoedavis/Documents/GitHub/EQI_final/functions/temperature_difference.R")

temp <-  temp_difference(raster = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/temperature_cold/montreal_cold.tif",
                         census = "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/DAs/montrealmetro.shp",
                         type = "cold",
                         shapefile_save_location =
                           "/Users/zoedavis/Documents/Projects/PHAC_EQI_2021/Data/Derived/montreal/montreal_cold.shp"
                           ) 


temp %>% 
    ggplot() +
    #geom_sf(aes(fill= as.numeric(d_DA_cold_temp_mean))) +
  geom_sf(aes(fill= as.numeric(temperature))) +
  scale_fill_gradientn(colors = terrain.colors(20)) +
  labs(fill = "difference in temp")

# d_DA_cold_temp_mean = difference between city mean and the DA temp
# mean_cold_temp = overall average temperature for the city
# cold_temperature  = average temperature within each DA

#temprtr
#mn_cold_
#d_DA___


```







